\textbf{Replies to Reviewers' Comments}

### 2023-10-14 Articles/2021-HarriganSpekkens/Submissions/FoP/response1.tex
Changes from Andrea

\textcolor{blue}{ For specific examples of ontological models, see \cite{Harrigan:2010}, sections 2.4.1--3. }

@@

\textbf{Authors:} TODO

@@

\begin{enumerate}
	\item \textbf{Reviewer 2:} Also, the technical results are correct. I would just simply say: quantum theory is a strongly non-Kolmogorovian probability calculus. Therefore, all attempts to embed it into a classical probability model will give place to problems. The core of the philosophical problem is to continue the pursuit of hidden-variables. They just don’t fit naturally. The PBR theorem is nothing but an expression of this inadequacy. But, of course, there is too much confusion out there. Taking into account that confusion, the results presented become quite relevant, Sincé they help to understand better what is wrong. For these reasons, I am willing to recommend this manuscript for publication.
	\vspace{2mm}

@@

	\textcolor{blue}{Let us denote $M(\Lambda)$ the space of all probability distributions (i.e. all probability measures) over the space of ontic states $\Lambda$. Let us denote $\mathcal{H}$ the Hilbert space of the corresponding quantum system and $M(\mathcal{H})$ the space of all mixed states (i.e. the space of positive semi-definite trace-one Hermitian operators). An ontological model, then, must give us an appropriate $\rho(\lambda) \in M(\Lambda)$ for every $\rho \in M(\mathcal{H})$ such that all predictions are satisfied. This includes those given by statistical mechanics and information theory, which depend on the von Neumann entropy $H_\mathcal{H} : M(\mathcal{H}) \to \mathbb{R}$. Given that $M(\Lambda)$ is the space of probability measures over $\Lambda$, it is natural to use the Shannon/Gibbs entropy function $H_\Lambda : M(\Lambda) \to \mathbb{R}$, as this is what is typically done in statistical mechanics.\footnote{Note that, while many functions called ``entropy'' exist, the Shannon entropy is the only indicator of variability that is continuous, monotonic and linear in probability.\cite{aop-phys-variability} It is also the one that provides the correct link to the thermodynamic entropy.} Since we want to be able to replicate the predictions of statistical mechanics and information theory, a very reasonable request is that these two entropies, calculated on the different representations, agree.
	}

@@

\textcolor{blue} {
	Now, the fact that quantum theory does not follow the rules of classical-Kolmogorov probability is nothing new[33]. However, this does not seem to be enough to clear confusion on the subject because, admittedly, we do use classical probability in the context of quantum mechanics. For example, we can imagine preparing the direction of spin based on a classical distribution. Furthermore, the output of a measurement is described by classical probability. The temptation, then, is to simply assume that all we need to do is just put the right transition probability between the two spaces, and we are done. This is exactly what cannot be done and what doesn't work. That is
	\begin{quote}
		\textbf{In the context of quantum mechanics, standard probability measures are allowed over preparations and measurement outcomes, but not over states.}
	\end{quote}
	In our approach, we exploited the fact that the entropy calculation uses the geometry of the inner product without invoking a probability of transition. That is, two pure states are orthogonal not because one cannot measure the first having prepared the second, but because their equal mixture raises the entropy by one bit. Therefore quantum mechanics is not simply defining a probability of transition between preparations and measurements: it is doing something more. While it is not in the scope of this article to articulate precisely what this ``more'' is, which will be the focus of future work, we hope that this additional insight may provide more precise guidance as to where classical probability calculus is appropriate and not in the context of quantum mechanics.	
}

\item \textbf{Reviewer 2:} - (3)    In page 5, in the conclusions, the authors mention that “quantum mixtures are non-classical”. But again, this is very well known. I think that the distinction between proper vs improper mixtures should be mentioned here. There is a nice discussion between Kirkpatrick and D’Espagnat on the arxives about this.
\vspace{2mm}

@@

\textcolor{blue} {The convex space formed by the mixtures is radically different[34]. There is also debate as whether proper and improper mixtures are distinguishable in the context of quantum mechanics[35-36].
}

@@

\textbf{Authors:} TODO.

@@

\textbf{Authors:} TODO.

>>>>

\textcolor{blue}{ For specific examples of ontological models, see [1], sections 2.4.1--3. }

@@

\textbf{Authors:} The referee highlights an interesting point in this comment. In the mentioned work Hance, Rarity and Ladyman criticize the assumptions on which Harrigan \& Spekkens classification rests on, and show that the dichotomy between $\psi$-ontic and $\psi$-epistemic models is artificial. In the revised manuscript we added a new paragraph at the end of section II mentioning this paper and how our argumentative strategy is complementary w.r.t.\ that employed by Hance, Rarity and Ladyman:

\textcolor{magenta}{After having introduced the relevant definitions of the HS ontological model framework, it should be noted that it has been subject to several criticisms, as already said in the previous section. For instance, Oldofredi \& Lopez in [10] analyzed the implicit assumptions made by Harrigan \& Spekkens in order to define their categorization, and showed that relevant interpretations of quantum theory---such as the statistical, relational and perspectival readings of the quantum formalism---cannot be classified according to the HS approach, given that their principles are in tension with those employed in [1]. Similarly, Hance, Rarity and Ladyman argued in [11] that the dichotomy ``ontic vs.\ epistemic'' proposed by HS is artificial, since there are cases in which the quantum state can represent both ontic properties of quantum systems as well as agents' knowledge of them. Thus, they claim, the definitions of $\psi$-ontic and $\psi$-epistemic introduced a few lines above do not capture the intuitive ideas with which physicists and philosophers alike use the terms ``ontic'' and ``epistemic''.}

\textcolor{magenta}{While we agree with the main messages of these critical responses to the HS framework, the argument offered in this essay follows a different and complementary strategy. In fact, instead of criticizing the assumptions used by Harrigan \& Spekkens, we consider them valid (for the sake of the argument) and derive a no-go theorem for $\psi$-ontic models directly from them, as we are going to show in the next section.}

The result obtained by Hance, Rarity and Ladyman is not in contrast with ours since both essays aim at showing the inadequacy of the HS in order to correctly classify interpretations of quantum mechanics. However, while it can be true that  the ``condition of non-overlapping can be interpreted as a sufficient, but not necessary for an ontic description'' as the referee correctly says, in our manuscript we consider the HS definitions as true in order to derive a no-go theorem for $\psi$-ontic directly from them.

@@

\begin{enumerate}
	\item \textbf{Reviewer 2:} Also, the technical results are correct. I would just simply say: quantum theory is a strongly non-Kolmogorovian probability calculus. Therefore, all attempts to embed it into a classical probability model will give place to problems. The core of the philosophical problem is to continue the pursuit of hidden-variables. They just don’t fit naturally. The PBR theorem is nothing but an expression of this inadequacy. But, of course, there is too much confusion out there. Taking into account that confusion, the results presented become quite relevant, Since they help to understand better what is wrong. For these reasons, I am willing to recommend this manuscript for publication.
	\vspace{2mm}

@@

	\textcolor{blue}{Let us denote $M(\Lambda)$ the space of all probability distributions (i.e. all probability measures) over the space of ontic states $\Lambda$. Let us denote $\mathcal{H}$ the Hilbert space of the corresponding quantum system and $M(\mathcal{H})$ the space of all mixed states (i.e. the space of positive semi-definite trace-one Hermitian operators). An ontological model, then, must give us an appropriate $\rho(\lambda) \in M(\Lambda)$ for every $\rho \in M(\mathcal{H})$ such that all predictions are satisfied. This includes those given by statistical mechanics and information theory, which depend on the von Neumann entropy $H_\mathcal{H} : M(\mathcal{H}) \to \mathbb{R}$. Given that $M(\Lambda)$ is the space of probability measures over $\Lambda$, it is natural to use the Shannon/Gibbs entropy function $H_\Lambda : M(\Lambda) \to \mathbb{R}$, as this is what is typically done in statistical mechanics. [Footnote: Note that, while many functions called ``entropy'' exist, the Shannon entropy is the only indicator of variability that is continuous, monotonic and linear in probability [30]. It is also the one that provides the correct link to the thermodynamic entropy.] Since we want to be able to replicate the predictions of statistical mechanics and information theory, a very reasonable request is that these two entropies, calculated on the different representations, agree.}

@@

\textcolor{blue} {
	Now, the fact that quantum theory does not follow the rules of classical-Kolmogorov probability is nothing new [33]. However, this does not seem to be enough to clear confusion on the subject because, admittedly, we do use classical probability in the context of quantum mechanics. For example, we can imagine preparing the direction of spin based on a classical distribution. Furthermore, the output of a measurement is described by classical probability. The temptation, then, is to simply assume that all we need to do is just put the right transition probability between the two spaces, and we are done. This is exactly what cannot be done and what doesn't work. That is
	\begin{quote}
		\textbf{In the context of quantum mechanics, standard probability measures are allowed over preparations and measurement outcomes, but not over states.}
	\end{quote}
	In our approach, we exploited the fact that the entropy calculation uses the geometry of the inner product without invoking a probability of transition. That is, two pure states are orthogonal not because one cannot measure the first having prepared the second, but because their equal mixture raises the entropy by one bit. Therefore quantum mechanics is not simply defining a probability of transition between preparations and measurements: it is doing something more. While it is not in the scope of this article to articulate precisely what this ``more'' is, which will be the focus of future work, we hope that this additional insight may provide more precise guidance as to where classical probability calculus is appropriate and not in the context of quantum mechanics.	
}

\item \textbf{Reviewer 2:} - (3)    In page 5, in the conclusions, the authors mention that “quantum mixtures are non-classical”. But again, this is very well known. I think that the distinction between proper vs improper mixtures should be mentioned here. There is a nice discussion between Kirkpatrick and D’Espagnat on the arXiv about this.
\vspace{2mm}

@@

\textcolor{blue} {The convex space formed by the mixtures is radically different [34]. There is also debate as whether proper and improper mixtures are distinguishable in the context of quantum mechanics [35-36].
}

@@

\textbf{Authors:} The reviewer raises a very important point that was not adequately discussed in the original version of the manuscript. In the revised version of the paper we added a new paragraph in the conclusion section saying that the HS framework should not be employed in order to assess the viability of interpretations of the quantum formalism---neither to claim that a certain reading of QM is empirically adequate, not to rule out a specific view:

\textcolor{magenta}{Referring to this, two final remarks are in order: on the one hand, our analysis entails an interesting consequence for the philosophical discussion concerning the interpretation of quantum theory. Indeed, many scholars use the PBR result to vindicate the empirical adequacy of a given framework---especially those in which the quantum state is somehow supposed to represent real physical objects---or to rule out a specific reading of the quantum formalism. Here we warn physicists and philosophers that no argument resting on Harrigan \& Spekkens approach should be employed in such metaphysical debates, given that no theory satisfying the requirements imposed by their categorization is able to reproduce \emph{all} the results of quantum mechanics.}

\textcolor{magenta}{To this regard, however, one may argue that our result cannot possibly be correct because there are actual examples of $\psi$-ontic models. To our knowledge, however, none of them is fully empirically equivalent to quantum theory, i.e.\ is able to reproduce all results from quantum statistical mechanics (e.g., derive the Boltzmann distribution from entropy maximization) or quantum information theory (e.g., derive the Holevo bound). Since the latter are derived from the more general framework of quantum mechanics, one cannot cherry pick which results a model has to reproduce to be a valid model of quantum theory: \emph{all} results must be correctly reproduced. In sum, we not only agree with Hance, Rarity and Ladyman in saying that HS definitions of $\psi$-ontic and $\psi$-epistemic models do not capture the actual meaning usually assigned to the terms ``ontic'' and ``epistemic'' as they are employed when interpreting quantum theories, but also say the following:}
\begin{quote}
\textcolor{magenta}{\textbf{If a certain interpretation of the quantum formalism is fully empirically equivalent to quantum mechanics, then it cannot be categorized as a $\psi$-ontic model.}}
\end{quote}
\noindent \textcolor{magenta}{In turn, the logical conjunction of our result and the PBR theorem entails that no interpretation of quantum theory can be correctly categorized according to the HS approach.}

Another general problem that we see---similar to that mentioned by referee 2---is that the notions of ``ontic'' and ``epistemic'' have always been liberally employed when interpreting quantum theories, but now it is commonly believed that HS provide a set of formal definitions capturing the meaning of such terms. In this way, many authors think to substantiate their views with the aid of HS classification. We believe that it will be very hard to deconstruct such an habit, however, our essay can be another step to achieve this goal. We explicitly discussed a paper of Hance, Rarity and Ladyman in section II, where this issue is openly addressed. For more details cf.\ the response to the fourth comment of referee 1.

@@

\textbf{Authors:} Again, we agree with the referee's comment. We firmly believe that the philosophical background of HS is problematic, as one of us argued in [10]. Following the suggestion we added another final paragraph to the conclusion section highlighting that the realist view about the quantum state---its ``onticity''---contained in $\psi$-ontic models is poorly defined:

\textcolor{magenta}{On the other hand, and we conclude, another philosophical issue affecting the definition of $\psi$-ontic model must be mentioned. In their essay Harrigan \& Spekkens require that if a model is $\psi$-ontic, the ontic state $\lambda$ is put into a  one-to-one correspondence with the quantum state, i.e.\ only one quantum state is compatible with the ontic state of the system. This fact, in turn, means that $\psi$ should be understood realistically, as an entity representing some real physical property of the system, or even the system itself as in $\psi$-complete models. However, as interestingly noted by Halvorson, realism towards the quantum state is highly problematic, since each representation relation proposed in literature does entail several conundrums that should be avoided---cf.\  \cite{Halvorson:2019} for details. To this regard, in fact, HS do not clearly spelled out what they mean when it is asserted that $\psi$ ``represents'' a real property of physical systems. Alternatively, to claim that $\psi$-ontic models $\lambda$ is uniquely represented by a single quantum state leaves many ontological questions unanswered: for instance it is not specified in which way $\psi$ represents the ontic state, is it a direct or indirect representation? If $\psi$ is a real object, as one would expect in $\psi$-complete models, is there a literal identification between $\psi$ and $\lambda$? To these and other issues Harrigan \& Spekkens do not reply.}

\noindent We did not included a more elaborated discussion of the latter point since the main focus of the paper is to prove the technical statement that $\psi$-ontic models cannot fully reproduce quantum theory. However, we will address the philosophical implications of our analysis in another philosophically oriented paper that is in preparation.  
\end{enumerate}
